### 2018-01-29 circuit-elements/paper.tex
Added polarization data plots

\input{header}
\physics
\begin{document}
\papertitle{Nonlinear Circuit Elements}
\paperauth{A}{Khesin}{1002442029}
\paperauth{P}{Zavyalova}{1002345036}
\paperdate{January 17, 2018}
\begin{paperabs}

	% a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a

	The relationships between current, potential, resistance, and temperature are of particular interest in nonlinear circuit components. In this experiment, these physical quantities were investigated for a thermistor and an incandescent light bulb. For the thermistor, resistance was found to decrease exponentially as temperature increased, following the relation \( R = \pars{1.0 \pm 0.1} \times 10^{13}e^{-T/\pars{24.7 \pm 0.3}} \). The relationship between potential and current of the light bulb was determined to obey a power law, \( V = \pars{0.067 \pm 0.002} I^{\pars{1.63 \pm 0.01}} \). Further, the exponential relational constant in Stefan-Boltzmann law was approximated to be \( 4.17 \pm 0.05 \).

@@

	The purpose of this experiment was to study the relationship between voltage and current for a thermistor and an incandescent lamp. Both of these circuit elements are nonlinear; for a thermistor, the resistance is a function of temperature and obeys the relation

@@

	The relation in (1) was investigated using a multimeter, a thermistor, a thermometer and a thermal flask for maintaining the thermistor at different temperatures, set up as shown below. Temperature of the thermistor was decreased by adding cold water to the thermal flask and stirring to ensure uniform temperature in the flask. \vspace{-0.8em}

@@

	In general, the tungsten filament in an incandescent light bulb obeys Ohm's law. As the temperature increases, however, the resistance of the metal increases and above a certain temperature becomes proportional to the absolute temperature. The actual relationship between \( V \) and \( I \) is investigated in this experiment by fitting obtained measurements of current and potential to

@@

	The current-potential data obtained for an incandescent light bulb was further used to determine the value of \( n \) in the Stefan-Boltzmann law,

@@

	Further, the value of \( n \) was obtained. Using the expression in (3) and assuming thermal equilibrium, \( \kappa \) and \( T \) were expressed in terms of \( I \) and \( V \). Additionally, by expressing \( V \) in terms of \( I \) as shown in (2), \( \kappa \) was written in terms of \( I \) and constants only.

@@

	The resistance of the thermistor at various temperatures was measured. The uncertainty, as specified by the manufacturer, was $ \pm (0.2\% + 1 $ LSD) for resistances in the $ (2 \si{\kilo\ohm}, 200 \si{\kilo\ohm}) $ range. The reading error of the thermometer was taken to be $ \pm 0.5 \si{\kelvin} $. \vspace{0.5em}

	\paperfig{thermistor}{\begin{papertable}{|[outer]I|[inner]C|[inner]U|[outer]}
			\paperoline\papertableindexheader&\textsc{Temperature}&\multicolumn{2}{c|[outer]}{\textsc{Resistance}}\\
			\papertablecheadersymbol{T$ $(\pm0.5$ $\si{\kelvin})}\papertableuheadersymbole{R$ $(\si{\kilo\ohm})}\\\paperiline
			\papertableindex\papertablecval{352.5}\papertableuval{6.68}{.02}\\\paperiline
			\papertableindex\papertablecval{349.0}\papertableuval{6.93}{.02}\\\paperiline
			\papertableindex\papertablecval{348.0}\papertableuval{7.54}{.03}\\\paperiline
			\papertableindex\papertablecval{346.0}\papertableuval{8.01}{.03}\\\paperiline
			\papertableindex\papertablecval{340.0}\papertableuval{9.95}{.03}\\\paperiline
			\papertableindex\papertablecval{338.0}\papertableuval{10.33}{.03}\\\paperiline
			\papertableindex\papertablecval{335.0}\papertableuval{11.76}{.03}\\\paperiline
			\papertableindex\papertablecval{334.5}\papertableuval{12.36}{.03}\\\paperiline
			\papertableindex\papertablecval{329.5}\papertableuval{14.45}{.04}\\\paperiline
			\papertableindex\papertablecval{327.5}\papertableuval{15.33}{.04}\\\paperiline
			\papertableindex\papertablecval{324.0}\papertableuval{17.23}{.04}\\\paperiline
			\papertableindex\papertablecval{321.0}\papertableuval{19.51}{.05}\\\paperiline
			\papertableindex\papertablecval{319.0}\papertableuval{21.7}{.1}\\\paperiline
			\papertableindex\papertablecval{317.0}\papertableuval{24.3}{.1}\\\paperiline
			\papertableindex\papertablecval{314.0}\papertableuval{26.1}{.2}\\\paperiline
			\papertableindex\papertablecval{309.5}\papertableuval{31.4}{.2}\\\paperiline
			\papertableindex\papertablecval{308.5}\papertableuval{33.7}{.2}\\\paperiline
			\papertableindex\papertablecval{307.5}\papertableuval{35.5}{.2}\\\paperiline
			\papertableindex\papertablecval{305.0}\papertableuval{38.0}{.2}\\\paperiline
			\papertableindex\papertablecval{304.0}\papertableuval{40.3}{.2}\\\paperiline
			\papertableindex\papertablecval{301.5}\papertableuval{44.3}{.2}\\\paperiline
			\papertableindex\papertablecval{299.5}\papertableuval{47.7}{.2}\\\paperiline
			\papertableindex\papertablecval{298.0}\papertableuval{51.6}{.2}\\\paperiline
			\papertableindex\papertablecval{296.0}\papertableuval{56.3}{.2}\\\paperiline
			\papertableindex\papertablecval{294.0}\papertableuval{61.3}{.2}\\\paperiline
			\papertableindex\papertablecval{292.5}\papertableuval{66.9}{.2}\\\paperiline
			\papertableindex\papertablecval{290.5}\papertableuval{73.8}{.2}\\\paperiline
			\papertableindex\papertablecval{286.0}\papertableuval{92.9}{.3}\\\paperiline
			\papertableindex\papertablecval{283.0}\papertableuval{104.7}{.3}\\\paperiline
			\papertableindex\papertablecval{278.0}\papertableuval{~126.2}{\parbox{\widthof{126.2~}}{.4}}\\\paperoline
		\end{papertable}\vspace{-1.5em}}
	{Thermistor's resistance at given temperatures. }\vspace{0.5em}

	Further, current and potential for a tungsten filament light bulb were measured and recorded. As specified by the manufacturer, the uncertainty in potential measurements was taken to be \( \pm (0.25\% + 1 \) LSD); current measurements carried an uncertainty of \( \pm (1\% + 3 \) LSD). \vspace{0.5em}

@@

	Upon propagating uncertainties in the obtained data using the method of quadratures, the relationship between the temperature and resistance of a thermistor was calculated using the expression in (1) as the model function. Least squares fitting of all data points yielded \( R_0 = (1.0 \pm 0.1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm 0.3) \si{\kelvin} \) as the fit parameters, with a reduced \( \chi ^ 2 \) value of \( 252.1710 \). The data, model, and residuals for this fit are shown below.

	\paperfig{thermistor-plot}{\pdf{Thermistor}}{Resistance and residuals plotted against the temperature in the thermal flask. The curve has the function \( (1.0 \pm 0.1) \times 10^{13}e^\frac{T}{\pars{24.7 \pm 0.3}} \) and an \( R^2 \) value of 0.992783. Error bars on both temperature and resistance were included but were too small to be visible.} \vspace{1em}

	A visualization of residuals clearly indicated that the data can be better modelled for greater temperatures. Indeed, least squares fitting performed without the data points corresponding to three lowest temperatures yielded a reduced \( \chi ^ 2 \) value of \( 181.5543 \). Removing more lower-temperature data points further improved reduced \( \chi ^ 2 \). 
	% Removing the six lower-temperature points led to a model with reduced \( \chi ^ 2 \) value of \( 130.1026 \).

	Further, a least squares fitting was performed on the data obtained for the incandescent light bulb, with the expression in (2) used as a model function. The formula was found to hold; the resulting parameters were \( k =  (0.067 \pm 0.002) \si{\volt\per\ampere} \) and \( a = (1.63 \pm 0.01) \). The value of reduced \( \chi ^ 2 \) obtained was \( 176.8952 \). The data, model, and residuals for this fit are shown below.

	\paperfig{bulb-plot}{\pdf{Bulb}}{Potential and residuals plotted against current through the incandescent bulb. The curve has the function \( \pars{0.067 \pm 0.002} I^{\pars{1.63 \pm 0.01}} \) with an \( R^2 \) value of \( 0.999532 \). Error bars on the data points were included but were too small to be visible.} \vspace{1em}

@@

	Upon obtaining the value of \( a \), the universal constant \( n \) in Stefan-Boltzmmann law could be approximated. Plugging \( a = (1.63 \pm 0.01) \) into (6) yielded \( 4.17 \pm 0.05 \) (uncertainty propagated using the method  of quadratures), which was close the actual value of \( n = 4\). Another constant appearing in Stefan-Boltzmann law, \( \epsilon \) (emissivity of the radiating body), differs depending on the material. \( \epsilon = 1 \) corresponds to a body that is a perfect absorber and emitter of radiant energy (i.e., a black body); \( \epsilon = 0 \) corresponds to a perfect reflector of radiant energy.

@@

		This experiment has investigated two nonlinear circuit elements, a thermistor and an incandescent tungsten filament light bulb. Resistance of the thermistor was found to be a function of the temperature, satisfying the exponential relation in (1) with \( R_0 = (1.0 \pm 0.1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm 0.3) \si{\kelvin} \).  For the light bulb, the relationship between potential and current was shown to follow a power law introduced in (2), with \( k =  (0.067 \pm 0.002) \si{\volt\per\ampere} \) and \( a = \pars{1.63 \pm 0.01} \). The most significant sources of error arose from potential uneven temperature of water in the thermal flask with the thermistor and inconsistencies in current and potential readings of the multimeter when switching the range of input signal.

		The above value of \( a \) was further used to approximate the power \( n \) of \( T \) in Stefan-Boltzmann law; \( n \) was found to be \( 4.17 \pm 0.05 \), reasonably close to the actual value of \( n = 4 \).

@@

\papersource{}

>>>>

\input{header}
\physics
\begin{document}
\papertitle{Nonlinear Circuit Elements}
\paperauth{A}{Khesin}{1002442029}
\paperauth{P}{Zavyalova}{1002345036}
\paperdate{January 17, 2018}
\begin{paperabs}
	The relationships between current, potential, resistance, and temperature are of particular interest in nonlinear circuit components. In this experiment, these physical quantities were investigated for a thermistor and an incandescent light bulb. For the thermistor, resistance was found to decrease exponentially as temperature increased, following the relation \( R = \pars{1.0 \pm .1} \times 10^{13}e^{-T/\pars{24.7 \pm .3}} \). The relationship between potential and current of the light bulb was determined to obey a power law, \( V = \pars{0.067 \pm .002} I^{\pars{1.63 \pm .01}} \). Further, the exponential relational constant in Stefan-Boltzmann law was found to be $\pars{4.17\pm.05}$.

@@

	The purpose of this experiment was to study the relationship between voltage and current for a thermistor and an incandescent lamp. These circuit elements are nonlinear; for a thermistor, resistance is a function of temperature and obeys the following relation.

@@

	The relation in (1) was investigated using a multimeter, a thermistor, a thermometer and a thermal flask for maintaining the thermistor at different temperatures, set up as shown below. Temperature of the thermistor was decreased by adding cold water to the thermal flask and stirring to ensure it was uniform. \vspace{-0.8em}

@@

	In general, the tungsten filament in an incandescent light bulb obeys Ohm's law. As the temperature increases, however, the resistance of the metal increases and above a certain temperature becomes proportional to the absolute temperature. The actual relationship between \( V \) and \( I \) was investigated in this experiment by fitting obtained measurements of current and potential to the following.

@@

	The current-potential data obtained for an incandescent light bulb was further used to determine the value of \( n \) in the Stefan-Boltzmann law.

@@

	Using the expression in (3) and assuming thermal equilibrium, \( \kappa \) and \( T \) were expressed in terms of \( I \) and \( V \). Additionally, by expressing \( V \) in terms of \( I \) as shown in (2), \( \kappa \) was written in terms of \( I \) and constants only.

@@

	The resistance of the thermistor at various temperatures was measured. The uncertainty, as specified by the manufacturer, was $ \pm (.2\% + 1 $ LSD) for resistances in the $ (2 \si{\kilo\ohm}, 200 \si{\kilo\ohm}) $ range. The reading error of the thermometer was taken to be $ \pm .5 \si{\kelvin} $. \vspace{0.5em}

	\paperfig{thermistor}{\begin{papertable}{|[outer]I|[inner]C|[inner]U|[outer]}
			\paperoline\papertableindexheader&\textsc{Temperature}&\multicolumn{2}{c|[outer]}{\textsc{Resistance}}\\
			\papertablecheadersymbol{T$ $(\pm.5$ $\si{\kelvin})}\papertableuheadersymbole{R$ $(\si{\kilo\ohm})}\\\paperiline
			\papertableindex\papertablecval{352.5}\papertableuval{6.68}{.02}\\\paperiline
			\papertableindex\papertablecval{349.0}\papertableuval{6.93}{.02}\\\paperiline
			\papertableindex\papertablecval{348.0}\papertableuval{7.54}{.03}\\\paperiline
			\papertableindex\papertablecval{346.0}\papertableuval{8.01}{.03}\\\paperiline
			\papertableindex\papertablecval{340.0}\papertableuval{9.95}{.03}\\\paperiline
			\papertableindex\papertablecval{338.0}\papertableuval{10.33}{.03}\\\paperiline
			\papertableindex\papertablecval{335.0}\papertableuval{11.76}{.03}\\\paperiline
			\papertableindex\papertablecval{334.5}\papertableuval{12.36}{.03}\\\paperiline
			\papertableindex\papertablecval{329.5}\papertableuval{14.45}{.04}\\\paperiline
			\papertableindex\papertablecval{327.5}\papertableuval{15.33}{.04}\\\paperiline
			\papertableindex\papertablecval{324.0}\papertableuval{17.23}{.04}\\\paperiline
			\papertableindex\papertablecval{321.0}\papertableuval{19.51}{.05}\\\paperiline
			\papertableindex\papertablecval{319.0}\papertableuval{21.7}{.1}\\\paperiline
			\papertableindex\papertablecval{317.0}\papertableuval{24.3}{.1}\\\paperiline
			\papertableindex\papertablecval{314.0}\papertableuval{26.1}{.2}\\\paperiline
			\papertableindex\papertablecval{309.5}\papertableuval{31.4}{.2}\\\paperiline
			\papertableindex\papertablecval{308.5}\papertableuval{33.7}{.2}\\\paperiline
			\papertableindex\papertablecval{307.5}\papertableuval{35.5}{.2}\\\paperiline
			\papertableindex\papertablecval{305.0}\papertableuval{38.0}{.2}\\\paperiline
			\papertableindex\papertablecval{304.0}\papertableuval{40.3}{.2}\\\paperiline
			\papertableindex\papertablecval{301.5}\papertableuval{44.3}{.2}\\\paperiline
			\papertableindex\papertablecval{299.5}\papertableuval{47.7}{.2}\\\paperiline
			\papertableindex\papertablecval{298.0}\papertableuval{51.6}{.2}\\\paperiline
			\papertableindex\papertablecval{296.0}\papertableuval{56.3}{.2}\\\paperiline
			\papertableindex\papertablecval{294.0}\papertableuval{61.3}{.2}\\\paperiline
			\papertableindex\papertablecval{292.5}\papertableuval{66.9}{.2}\\\paperiline
			\papertableindex\papertablecval{290.5}\papertableuval{73.8}{.2}\\\paperiline
			\papertableindex\papertablecval{286.0}\papertableuval{92.9}{.3}\\\paperiline
			\papertableindex\papertablecval{283.0}\papertableuval{104.7}{.3}\\\paperiline
			\papertableindex\papertablecval{278.0}\papertableuval{~126.2}{\parbox{\widthof{126.2~}}{.4}}\\\paperoline
		\end{papertable}\vspace{-1.5em}}
	{Thermistor's resistance at given temperatures. }\vspace{0.5em}

	Further, current and potential for a tungsten filament light bulb were measured and recorded. As specified by the manufacturer, the uncertainty in potential measurements was taken to be \( \pm (.25\% + 1 \) LSD); current measurements carried an uncertainty of \( \pm (1\% + 3 \) LSD). \vspace{0.5em}

@@

	Upon propagating uncertainties in the obtained data using the method of quadratures, the relationship between the temperature and resistance of a thermistor was calculated using the expression in (1) as the model function. Least squares fitting of all data points yielded \( R_0 = (1.0 \pm .1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm .3) \si{\kelvin} \) as the fit parameters, with a reduced \( \chi ^ 2 \) value of \( 252.1710 \). The data, fit, and residuals are shown below.

	A visualization of residuals  indicated that the data can be better modeled at higher temperatures. Least squares fitting without the data points corresponding to three lowest temperatures yielded a reduced \( \chi ^ 2 \) value of \( 181.5543 \). Removing more lower-temperature points improved it further.

	\paperfig{thermistor-plot}{\pdf{Thermistor}}{Resistance and residuals plotted against the temperature in the thermal flask. The curve has the function \( (1.0 \pm .1) \times 10^{13}e^{T/\pars{24.7 \pm .3}} \) and an \( R^2 \) value of 0.992783. Error bars on both temperature and resistance were included but were too small to be visible.}

	Further, a least squares fitting was performed on the data obtained for the incandescent light bulb, with the expression in (2) used as a model function. The formula was found to hold; the resulting parameters were \( k =  (0.067 \pm .002) \si{\volt\per\ampere} \) and \( a = (1.63 \pm .01) \). The value of reduced \( \chi ^ 2 \) obtained was \( 176.8952 \). The data, model, and residuals for this fit are shown below.

	\paperfig{bulb-plot}{\pdf{Bulb}}{Potential and residuals plotted against current through the incandescent bulb. The curve has the function \( \pars{0.067 \pm .002} I^{\pars{1.63 \pm .01}} \) with an \( R^2 \) value of \( 0.999532 \). Error bars on the data points were included but were too small to be visible.}

@@

	Upon obtaining the value of \( a \), the universal constant \( n \) in Stefan-Boltzmmann law could be approximated. Plugging \( a = (1.63 \pm .01) \) into (6) yielded $\pars{4.17\pm.05}$ (uncertainty propagated using the method  of quadratures), which was close the actual value of \( n = 4\). Another constant appearing in Stefan-Boltzmann law, \( \epsilon \) (emissivity of the radiating body), differs depending on the material. \( \epsilon = 1 \) corresponds to a body that is a perfect absorber and emitter of radiant energy (i.e., a black body); \( \epsilon = 0 \) corresponds to a perfect reflector of radiant energy.

@@

		This experiment has investigated two nonlinear circuit elements, a thermistor and an incandescent tungsten filament light bulb.\vfill\phantom{.}\columnbreak\noindent Resistance of the thermistor was found to be a function of the temperature, satisfying the exponential relation in (1) with \( R_0 = (1.0 \pm .1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm .3) \si{\kelvin} \).  For the light bulb, the relationship between potential and current was shown to follow a power law introduced in (2), with \( k =  (0.067 \pm .002) \si{\volt\per\ampere} \) and \( a = \pars{1.63 \pm .01} \).

		The most significant sources of error arose from potentially uneven temperature of water in the thermal flask with the thermistor where there could have been momentary hot or cold pockets and inconsistencies in current and potential readings of the multimeter when switching the range of input signal on the multimeters.

		The above value of \( a \) was further used to approximate the power \( n \) of \( T \) in Stefan-Boltzmann law; \( n \) was found to be \( 4.17 \pm .05 \), reasonably close to the actual value of \( n = 4 \).

@@

### 2018-01-29 polarization/paper.tex
Added polarization data plots

	Finally, intensity vs. position data was collected for the experimental setup measuring

>>>>

	Finally, intensity vs. position data was collected for the experimental setup investigating polarization by reflection. Uncertainties were taken to be \textbf{TODO}. 
	\paperfig{Exercise3-Data}{\pdf{exercise3-data}}{Intensity of reflected ray as a function of position of the rotary motion sensor. Error bars are not included \textbf{TODO} explain why.}
